---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# larus

<!-- badges: start -->
<!-- badges: end -->

The goal of larus is to 

Provide data example

Identify trips:
plot_check
trip_number_sequence
plot_trips
add_gaps

Identify is trips are valid by:
Calculating gaps between locations
Evaluate if trips are valid by checking thresholds (e.g. minimum duration away from the colony)

Recalculate foraging parameters if the bird relocate with:
create_buffer
from_df_to_sf
over 
add_trip_number



lenght paths requeire at least 3 locations per forafging trip
exclude trips where the bird was constantly moving from central location

recalculate foraging trips changing central locations

Interpolate without:
overestimating resting locations at night (make a filter of trips <24 hrs)
add a classification trip good for calculating foraging parameters
trip good for interpolation 


## Installation

You can install the development version of larus from [GitHub](https://github.com/) with:

```{r,eval=FALSE}
# install.packages("devtools")
devtools::install_github("MiriamLL/larus")
```

# ~~~~~~~~~~~~~~~~~

# Intro

There is two key issues in determining the foraging trips in GSMs attached to gulls:

1. The effect of the battery changes
2. The gulls just do whatever they want. 

# ~~~~~~~~~~~~~~~~~

## 1. The effect of the battery

```{r}
load("~/03Academico/02Proyectos-Postdoc/2025/1Programming/2Packages/larus/data/LALI_example.rda")
```

```{r}
LALI_01locs<-LALI03_02locs
```

When studying tracking locations from GSM data, especially in the case of birds like gulls, it’s common to encounter gaps in the data due to factors such as loss of signal, battery limitations, or movement through areas with poor GSM coverage. Whether to interpolate the data depends on your research goals and the nature of the gaps.

Here’s a breakdown of considerations to help you decide:

- Short Gaps: If the gaps are short (e.g., a few minutes to an hour), interpolation might be justifiable, especially if you assume that the bird's movement is relatively constant during the gap. For example, linear interpolation can estimate a reasonable position based on previous and subsequent data points.
- Long Gaps: If the gaps are long (e.g., several hours or days), interpolation might introduce unrealistic results, especially if the bird’s movement during that time period was unknown. Interpolating long gaps could lead to misleading conclusions about the bird’s actual path or behavior.
If your study’s analysis can handle missing data or gaps (e.g., for high-level migration patterns), you might consider leaving the gaps unfilled. This avoids the risk of introducing erroneous data but may reduce the overall precision.
If your analysis is based on trajectory or movement behavior, you could use gap-filling approaches like predicting the bird’s behavior based on its last known location and surrounding environmental context (such as weather, time of day, etc.).

```{r}
library(scales)
```

```{r,eval=FALSE}
ggplot(LALI_01locs, aes(x=daytime, y=battery.charge.percent)) +
  geom_line() +
  scale_x_datetime(labels = date_format("%b"),date_breaks = "1 month")
```


# ~~~~~~~~~~~~~~~~~

## 2. Identify trips

### Step 1: Subset month

```{r}
library(tidyverse)
```

```{r}
LALI_02locs<-as.data.frame(LALI03_02locs)
```

```{r}
LALI_08ago_01locs<-LALI_02locs%>%
  filter(month=="08")
```

### Step 2: Check central location

#### function: plot_check

```{r}
plot_check<-function(my_locs=my_locs){
  ggplot()+
  geom_point(data = my_locs, aes(x=Longitude, y = Latitude),
             color='black',size = 0.8,alpha=0.4)+
  geom_point(data=Gaviota_Island, aes(x=Longitude, y=Latitude),color='red',shape=17, size=5)+
  geom_density_2d_filled(data = my_locs, aes(x = Longitude, y = Latitude),alpha = 0.5)+
  theme_bw()+
  theme(legend.position = 'none')+
  ggtitle('Check that the triangle is close to the highest locations concentrations, \nthis should be your central location,  \nif the central location is shifted  \nrecalculate from central location')}
```

```{r}
Gaviota_Island<-data.frame(Longitude=-110.33979846296234,Latitude=24.28728834326802)
```

```{r}
plot_check(my_locs=LALI_08ago_01locs)
```

### Step 3: remove central locations

```{r}
LALI_08ago_02outside<-LALI_08ago_01locs %>%
  filter(inside=='outside_central')
```

### Step 4: trip_id

#### function: trip_number_sequence

```{r}
continue_trip_sequence<-function(my_previous){
  print(last(my_previous))
  trip_number_sequence<-substr(last(my_previous), start = 6, stop = 10)
  trip_number_sequence<-as.numeric(trip_number_sequence)
  print(trip_number_sequence)
return(trip_number_sequence)}
```

```{r}
LALI_07jul_params<-as.data.frame(trip_id<-"trip_00405")
```

```{r}
trip_number_sequence<-continue_trip_sequence(my_previous=LALI_07jul_params$trip_id)
```

```{r}
LALI_08ago_03trips<-LALI_08ago_02outside %>%
  mutate(num_seq=as.numeric(num_seq))%>%
  mutate(trip_number = (cumsum(c(1L, diff(num_seq)) !=   1L)))%>%
  mutate(trip_number = trip_number +1 + trip_number_sequence)%>%
  mutate(trip_number = stringr::str_pad(trip_number,  5, pad = "0"))%>%
  mutate(trip_number = paste0("trip_", trip_number))
```

### Step 5: check trips

function: plot_trips

```{r}
plot_trips<-function(my_locs=my_locs){
  ggplot()+
  geom_point(data = my_locs, aes(x=Longitude, y = Latitude,color=trip_number),
            size = 0.8,alpha=0.4)+
  geom_point(data=Gaviota_Island, aes(x=Longitude, y=Latitude),color='red',shape=17, size=5)+
  geom_density_2d_filled(data = my_locs, aes(x = Longitude, y = Latitude),alpha = 0.1)+
  theme_bw()+
  theme(legend.position = 'none')+
  ggtitle('Here excludes all nest locations \n See if there are patterns ')}
```

```{r}
plot_trips(my_locs=LALI_08ago_03trips)
```

### Step 6: check parameters

```{r}
library(sula)
```

```{r}
LALI_08ago_04params<-sula::calcular_duracion(GPS_data=LALI_08ago_03trips, 
                        col_diahora='daytime',
                        formato =  "%Y-%m-%d %H:%M:%S",
                        unidades="hours",
                        separador="trip_number")
```

Define trip_month_id in case there are outliers and the locations need to be subsetted

```{r}
LALI_08ago_04params<-LALI_08ago_04params %>%
  mutate(trip_month_id=paste0("ago_a",trip_id))%>%
  mutate(central_loc='colony')
```

### Step 7: add gaps

#### function: add_gaps

```{r}
add_gaps<-function(my_params=my_params,my_locs=my_locs){
  
  my_gaps<-my_locs %>%
  group_by(trip_number)%>%
  summarise(min_gap=min(Gaps_time),
            max_gap=max(Gaps_time))%>%
  rename(trip_id=trip_number)

    this_params<-my_params %>%
    left_join(my_gaps,by='trip_id')%>%
    select(trip_id,trip_month_id,central_loc,trip_start,trip_end,duration,min_gap,max_gap)
  
  return(this_params)
}
```

```{r}
LALI_08ago_05params<-add_gaps(my_params = LALI_08ago_04params, my_locs = LALI_08ago_03trips)
head(LALI_08ago_05params)
```


# ~~~~~~~~~~~~~~~~~

## 3. Gulls just do whatever they want

### Step 8: check outliers

```{r}
range(LALI_08ago_04params$duration)
```

```{r}
hist(LALI_08ago_04params$duration)
```

Check if the big outliers are because of gaps in the intervals.
In this case, the gap is not so large (9 to 20 minutes).
It is likely that the bird just change central location! 

```{r}
LALI_08ago_05params %>%
  filter(duration > 24)
```

#### identify outlier trips

Sometimes the gulls are free of responsibilities and decide to leave for very long periods.

Identify ids 

```{r}
LALI_08ago_tripid<-unique(LALI_08ago_05params %>%
  filter(duration > 40))$trip_id
LALI_08ago_tripid
```

Subset locations

```{r}
LALI_08ago_06outlierlocs<-LALI_08ago_03trips %>%
  filter(trip_number %in% LALI_08ago_tripid)%>%
  relocate(ID,daytime,dt,Gaps_time)
head(LALI_08ago_06outlierlocs)
```

Sometimes its possible to identify a new central location

```{r}
central_location<-data.frame(Longitude=-110.384,Latitude=24.108)
```

```{r}
plot_trips(LALI_08ago_06outlierlocs)+
  geom_point(data=central_location, aes(x=Longitude, y=Latitude),color='blue',shape=17, size=5)
```

### Step 9: Create buffer

#### function: create_buffer

```{r}
create_buffer<-function(central_point=central_point, buffer_km=buffer_km){
  central_spatial<- sp::SpatialPoints(cbind(central_point$Longitude,central_point$Latitude)) 
  sp::proj4string(central_spatial)= sp::CRS("+init=epsg:4326") 
  central_spatial <- sp::spTransform(central_spatial, sp::CRS("+init=epsg:4326"))
  central_spatial<-sf::st_as_sf(central_spatial)
  buffer_dist<-buffer_km*1000
  central_buffer<-sf::st_buffer(central_spatial, buffer_dist)
  return(central_buffer)
  }
```

```{r}
central_buffer<-create_buffer(central_point=central_location,buffer_km=0.3)
```

### Step 10: From df to sf

#### function: from_df_to_sf

```{r}
from_df_to_st<-function(my_df){
  my_points <- my_df
  sp::coordinates(my_points) <- ~Longitude + Latitude
  sp::proj4string(my_points) = sp::CRS("+init=epsg:4326")
  my_sf<-sf::st_as_sf(my_points)
  return(my_sf)
}
```

```{r}
LALI_08ago_09points<-from_df_to_st(LALI_08ago_06outlierlocs)
```

### Step 11: Over

```{r}
over<-function(this_location=this_location,this_buffer=this_buffer){
  locations<-sf::st_as_sf(this_location)
  location_over<-sapply(sf::st_intersects(locations,this_buffer), 
                        function(z) if (length(z)==0) NA_integer_ else z[1])
  return(location_over)}
```

```{r}
LALI_08ago_10over<- over(LALI_08ago_09points,central_buffer)
```

### Step 12: Add_trip_number

```{r}
add_trip_number<-function(my_over,my_df,my_previous){
  
  my_df$central_location <- my_over 
  
  my_df<-my_df %>%
    mutate(inside=case_when(central_location == 1 ~ 'inside_central',TRUE ~ 'outside_central'))
  
  my_df %>%
    group_by(ID,inside)%>%
    count()%>%
    pivot_wider(names_from = inside, values_from = n)
  
  my_outside<-my_df %>%
    filter(inside=='outside_central')
  
  last(my_previous$trip_id)
  
  trip_number_sequence<-substr(last(my_previous$trip_id), start = 6, stop = 10)
  trip_number_sequence<-as.numeric(trip_number_sequence)
  trip_number_sequence

  my_trips<-my_outside %>%
    mutate(num_seq=as.numeric(num_seq))%>%
    mutate(trip_number = (cumsum(c(1L, diff(num_seq)) !=   1L)))%>%
    mutate(trip_number = trip_number + 1 + trip_number_sequence)%>%
    mutate(trip_number = stringr::str_pad(trip_number,  5, pad = "0"))%>%
    mutate(trip_number = paste0("trip_", trip_number))
  
  return(my_trips)
}
```

```{r}
LALI_08ago_11trips<-add_trip_number(my_df=LALI_08ago_06outlierlocs,
                                     my_over=LALI_08ago_10over,
                                     my_previous=LALI_08ago_05params)
```

### Step 13: check parameters

### params

```{r}
LALI_08ago_12params<-sula::calcular_duracion(GPS_data=LALI_08ago_11trips, 
                        col_diahora='daytime',
                        formato =  "%Y-%m-%d %H:%M:%S",
                        unidades="hours",
                        separador="trip_number")
```

```{r}
LALI_08ago_12params<-LALI_08ago_12params %>%
  mutate(trip_month_id=paste0("ago_b_",trip_id))%>%
  mutate(central_loc='south_of_colony')
```

### add gaps

```{r}
LALI_08ago_13params<-add_gaps(my_params = LALI_08ago_12params, my_locs = LALI_08ago_11trips)
LALI_08ago_13params
```


```{r}
range(LALI_08ago_13params$duration)
```

```{r}
hist(LALI_08ago_13params$duration)
```

```{r}
plot_trips(my_locs=LALI_08ago_11trips)+
  geom_point(data=central_location, aes(x=Longitude, y=Latitude),color='blue',shape=17, size=5)
```


# merge and check 

## params

```{r}
LALI_08ago_06tripid<-unique(LALI_08ago_05params %>%
  filter(duration > 24))$trip_id
```

```{r}
LALI_08ago_params<-rbind(LALI_08ago_05params %>% filter(!trip_id %in% LALI_08ago_06tripid),
                         LALI_08ago_13params)
```

## locs

```{r}
LALI_08ago_locs<-rbind(LALI_08ago_03trips %>% 
                         filter(!trip_number %in% LALI_08ago_06tripid)%>%
                         select(ID,daytime,season,month,
                           Longitude,Latitude,countday,
                           Hour,Day_or_night,central_base,inside,num_seq,
                           ID.y,dt,Location_gaps,Battery,Gaps_time,battery.charge.percent,
                           trip_number),
                         
                         LALI_08ago_11trips %>%
                            select(ID,daytime,season,month,
                           Longitude,Latitude,countday,
                           Hour,Day_or_night,central_base,inside,num_seq,
                           ID.y,dt,Location_gaps,Battery,Gaps_time,battery.charge.percent,
                           trip_number))
```

```{r}
length(unique(LALI_08ago_locs$trip_number))
length(unique(LALI_08ago_params$trip_id))
```


# ~~~~~~~~~~~~~~~~~~~~~

# Straight forward


```{r}
LALI_01locs<-LALI_02locs%>%
  filter(month=="09")
```

```{r}
plot_check(my_locs=LALI_01locs)
```

## trip_id

```{r}
LALI_02outside<-LALI_01locs %>%
  filter(inside=='outside_central')
```

```{r}
trip_number_sequence<-continue_trip_sequence(my_previous=LALI_07jul_params$trip_id)
```

```{r}
LALI_03trips<-LALI_02outside %>%
  mutate(num_seq=as.numeric(num_seq))%>%
  mutate(trip_number = (cumsum(c(1L, diff(num_seq)) !=   1L)))%>%
  mutate(trip_number = trip_number +1 + trip_number_sequence)%>%
  mutate(trip_number = stringr::str_pad(trip_number,  5, pad = "0"))%>%
  mutate(trip_number = paste0("trip_", trip_number))
```

```{r}
plot_trips(my_locs=LALI_03trips)
```

## params

```{r}
LALI_04params<-sula::calcular_duracion(GPS_data=LALI_03trips, 
                        col_diahora='daytime',
                        formato =  "%Y-%m-%d %H:%M:%S",
                        unidades="hours",
                        separador="trip_number")
```

```{r}
range(LALI_04params$duration)
```

```{r}
hist(LALI_04params$duration)
```

```{r}
LALI_04params<-LALI_04params %>%
  mutate(trip_month_id=paste0("ago_a",trip_id))%>%
  mutate(central_loc='colony')
```

## gaps

```{r}
LALI_05params<-add_gaps(my_params = LALI_04params, my_locs = LALI_03trips)
head(LALI_05params)
```

## checks

```{r}
LALI_locs<-LALI_03trips
LALI_params<-  LALI_05params
```

```{r}
length(unique(LALI_params$trip_id))
length(unique(LALI_locs$trip_number))
```



# ~~~~~~~~~~~~~~~~~~~~~~~~

# Classify 

## Params

```{r}
Params<-LALI_params
```

```{r}
Params_01class<-Params %>%
  
  mutate(params_analyses=case_when(duration >= 24 ~ 'exclude_from_analyses',
                                     duration <= 24 ~ 'include_in_analyses',
                                TRUE ~ "other"))%>%
  
  mutate(resolution=case_when(max_gap >= 60 ~ 'low_resolution_gaps_more_60_mins',
                                max_gap <= 60 ~ 'ok_resolution_gaps_less_60_mins',
                                TRUE ~ "other"))%>%
  
  mutate(interpolation = case_when(duration >= 24 & max_gap >= 60 ~ 'dont_interpolate',
                                   duration >= 24 ~ 'dont_interpolate',
                                   max_gap >= 60 ~ 'dont_interpolate',
                                   duration <= 24 & max_gap <= 60 ~ 'interpolate',
                                   TRUE ~ "other"))%>%
  
  mutate(trip_size = case_when(duration >= 0.5 ~ 'trip_longer_than_30mins',
                               duration <= 0.5 ~ 'trip_shorter_than_30mins',
                               TRUE ~ "other"))
  
```


## check criteria

Foraging trips were only considered when the animal was more than 1 km away from the colony and lasting longer than 30 min (Shaffer et al. 2017, Guerra et al. 2022). 

```{r}
Params_01class %>%
   group_by(trip_size)%>%
   tally()
```

Candidates for interpolation

If trips had a gap of > 1hr and/or included overnight locations without a clear central location, they were excluded from the interpolation. 

```{r}
Params_01class %>%
  group_by(interpolation)%>%
  tally()
```

### Keep candidates

```{r}
Params_02class<-Params_01class %>%
  filter(trip_size=='trip_longer_than_30mins')
```


## Locs

### identify trips ids

```{r}
Locs_01analysesid<-unique(Params_02class %>%
  filter(params_analyses == "include_in_analyses"))$trip_id

Locs_02resolutionid<-unique(Params_02class %>%
  filter(resolution == "ok_resolution_gaps_less_60_mins"))$trip_id

Locs_03interpolateid<-unique(Params_02class %>%
  filter(interpolation == 'interpolate'))$trip_id

Locs_04sizeid<-unique(Params_02class %>%
 filter(trip_size == 'trip_longer_than_30mins'))$trip_id
```

```{r}
Locs_01class<-LALI_08ago_locs %>%
  mutate(params_analyses=case_when(trip_number %in% Locs_01analysesid~ "include_in_analyses",
                          TRUE ~ "exclude_from_analyses"))%>%
  
  mutate(resolution=case_when(trip_number %in% Locs_02resolutionid~ "ok_resolution_gaps_less_60_mins",
                          TRUE ~ "low_resolution_gaps_more_60_mins"))%>%
  
  mutate(interpolation=case_when(trip_number %in% Locs_03interpolateid ~ 'interpolate',
                          TRUE ~ "dont_interpolate"))%>%
  
  mutate(trip_size = case_when(trip_number %in% Locs_03interpolateid ~  'trip_longer_than_30mins',
                                 TRUE ~  'trip_shorter_than_30mins'))
```

# ~~~~~~~~~~~~~~~~~~~~~
# Test data

## gull_data

Contains data from one gull tagged at La Paz, Mexico

```{r,echo=FALSE,eval=FALSE}
#load packages
library(tidyverse)
#load data
LALIS_1807 <- read_csv("~/03Academico/01MS/01EnPreparacion/MS4 Lali/2023/02Jul/1Data/LALIS_1807.csv")
range(LALIS_1807$timestamp)
#rename columns
LALIS<-LALIS_1807 %>%
  filter(timestamp > '2023-04-20 00:00:00 UTC' & timestamp > '2023-04-27 00:00:00 UTC')
LALIS<-LALIS %>%
  select(timestamp,location.long,location.lat,individual.local.identifier)
LALIS$Longitude<-LALIS$location.long
LALIS$Latitude<-LALIS$location.lat
#identify trips
LALI02<-LALIS%>%
  filter(individual.local.identifier=='LALI02')
#devtools::install_github("MiriamLL/sula")
library(sula)
colony<-data.frame(Longitude=-110.338935,Latitude=24.287935)
LALIS_trips<-identificar_viajes(GPS_data=LALI02,
                        nest_loc=colony,
                        distancia_km=1)
LALIS_trips$at_sea<-LALIS_trips$trip
LALIS_atsea<-contar_viajes(GPS_data=LALIS_trips)
gull_data<-LALIS_atsea %>%
  select(individual.local.identifier,
         Longitude,Latitude,
         timestamp,
         at_sea,
         trip_number)%>%
  rename(ID=individual.local.identifier,
         DateTime=timestamp)
ggplot()+
  geom_point(data=gull_data, aes(x=Longitude,y=Latitude,color=trip_number))+
  theme_minimal()+
  theme(legend.position = 'none')
gull_data<-gull_data[1:1458,]
save(gull_data, file = "gull_data.RData", version = 2)
#usethis::use_r("gull_data")
devtools::document()
devtools::check()
```

```{r}
library(larus)
```

```{r}
gull_data<-gull_data
```

