---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# larus

<!-- badges: start -->
<!-- badges: end -->


## Installation

You can install the development version of larus from [GitHub](https://github.com/) with:

```{r,eval=FALSE}
# install.packages("devtools")
devtools::install_github("MiriamLL/larus")
```


# Intro

There is three key issues in determining the foraging trips in GSMs attached to gulls:

 1. Battery charge and gaps (intervals) in the data
 2. Identify foraging trips, classify is usable or not
 3. The gulls just do whatever they want. 

# ~~~~~~~~~~~~~~~~~

# 1. The effect of the battery

```{r,echo=FALSE,eval=FALSE}
load("~/03Academico/02Proyectos-Postdoc/2025/1Programming/2Packages/larus/data/LALI_example.rda")
GSM_battery<-GSM_battery %>%
  rename(battery_charge=battery.charge.percent,
         gaps_min=Gaps_time)%>%
  select(ID,daytime,season,month,battery_charge,gaps_min)
#Move to folder data
save(GSM_battery, file = "GSM_battery.RData", version = 2)
#Run in console, not Rmd
usethis::use_r("GSM_battery")
devtools::document()
devtools::check()
```

```{r}
library(larus)
```

```{r}
GSM_battery
```


```{r, message=FALSE}
library(scales)
library(ggplot2)
```

```{r,eval=FALSE}
ggplot(GSM_battery, aes(x=daytime, y=battery_charge)) +
  geom_line() +
  scale_x_datetime(labels = date_format("%b"),date_breaks = "1 month")
```

## Battery levels classification

Considerations:
Optimal battery was considered > 75% battery 
Good < 75% battery 
Poor < 50% battery 
Low  < 25% battery 

```{r, message=FALSE}
library(tidyverse)
```

```{r}
GSM_battery <- GSM_battery%>%
  mutate(battery_class = 
         case_when(is.na(battery_charge) ~ 'U',
                   battery_charge >= 75 ~ '1Optimal (>75%)',
                   battery_charge <= 75 & battery_charge > 50 ~ '2Good (>50%)',
                   battery_charge <= 50 & battery_charge > 25 ~ '3Poor (>25%)',
                   battery_charge <= 25 ~ '4Low (<25%)',
                   TRUE ~ 'Check'))
```

```{r}
Battery_classes <-GSM_battery %>%
  group_by(battery_class)%>%
  tally()%>%
  mutate(total=sum(n))%>%
  mutate(prop=n*100/total)
Battery_classes
```

## Gaps and battery relations

When studying tracking locations from GSM data, especially in the case of birds like gulls, it’s common to encounter gaps in the data due to factors such as loss of signal, battery limitations, or movement through areas with poor GSM coverage. 

Considerations:
Optimal gaps < 5 minutes 
Good < 15 minutes
Average < 30 minutes
Poor < 2 hours (120 minutes)
Low  > 24 hours (1400 minutes)

```{r}
Battery_gaps <- GSM_battery %>%
  mutate(gaps_class = 
           case_when(is.na(gaps_min) ~ '8U',
                     gaps_min <= 5 ~ '1Optimal (<5mins)',
                     gaps_min >= 5 & gaps_min < 15 ~ '2Good (<15mins)',
                     gaps_min >= 15 & gaps_min < 30 ~ '3Average (<30mins)',
                     gaps_min >= 30 & gaps_min < 60 ~ '4Poor (<1hr)',
                     gaps_min >= 60 & gaps_min < 1400 ~ '5Low (>1hr)',
                     gaps_min >= 1400 ~ '6Low (onceAday)',
                     TRUE ~ '7Check'))
```

```{r}
Battery_gaps %>%
  group_by(gaps_class)%>%
  tally()
```

## Compare 

```{r}
library(patchwork)
```

### Battery and gaps per season

```{r, fig.height=8, fig.width=7}
Battery_gaps%>%
  group_by(season,battery_class)%>%
  tally()%>%
  mutate(total=sum(n))%>%
  mutate(prop=n*100/total) %>%
  
ggplot(aes(fill=battery_class, y=prop, x=season)) + 
  geom_bar(stat="identity")+
  scale_fill_manual(values=c('#d7263d','#f46036','#2e294e'))+
  theme_classic()+
  scale_y_continuous(expand = c(0,0)) +
  
Battery_gaps %>%
  group_by(season,gaps_class)%>%
  tally()%>%
  mutate(total=sum(n))%>%
  mutate(prop=n*100/total)%>%
  filter(gaps_class!='8U') %>% #expect first position to be empty
  
ggplot(aes(fill=gaps_class, y=prop, x=season)) +
  geom_bar(stat="identity") +
  scale_fill_manual(values=c('#d7263d','#f46036','#ffc857','#c5d86d','#1b998b','#2e294e'))+
  theme_classic()+
  scale_y_continuous(expand = c(0,0)) +
  
plot_layout(ncol = 1)
```

### Battery and gaps per month

```{r, fig.height=8}
Battery_gaps%>%
  group_by(month,battery_class)%>%
  tally()%>%
  mutate(total=sum(n))%>%
  mutate(prop=n*100/total) %>%
  
ggplot(aes(fill=battery_class, y=prop, x=month)) + 
  geom_bar(stat="identity")+
  scale_fill_manual(values=c('#d7263d','#f46036','#2e294e'))+
  theme_classic()+
  scale_y_continuous(expand = c(0,0)) +
  
Battery_gaps %>%
  group_by(month,gaps_class)%>%
  tally()%>%
  mutate(total=sum(n))%>%
  mutate(prop=n*100/total)%>%
  filter(gaps_class!='8U') %>% #expect first position to be empty
  
ggplot(aes(fill=gaps_class, y=prop, x=month)) +
  geom_bar(stat="identity") +
  scale_fill_manual(values=c('#d7263d','#f46036','#ffc857','#c5d86d','#1b998b','#2e294e'))+
  theme_classic()+
  scale_y_continuous(expand = c(0,0)) +
  
plot_layout(ncol = 1)
```

# ~~~~~~~~~~~~~~~~~

## 2. Identify trips

### Step 1: Subset month

```{r,echo=FALSE,eval=FALSE}
load("~/03Academico/02Proyectos-Postdoc/2025/1Programming/2Packages/larus/data/LALI_example.rda")
GSM_locs<-LALI03_02locs
GSM_locs<-GSM_locs %>%
  rename(battery_charge=battery.charge.percent,
         gaps_min=Gaps_time)%>%
  select(num_seq,
         ID,
         daytime,season,month,
         Longitude,Latitude,
         inside,central_base,
         battery_charge,gaps_min)
#Move to folder data
save(GSM_locs, file = "GSM_locs.RData", version = 2)
#Run in console, not Rmd
usethis::use_r("GSM_locs")
devtools::document()
devtools::check()
```

```{r}
head(GSM_locs)
```

```{r, message=FALSE}
library(tidyverse)
```

```{r}
This_month<-'08'
This_month_text<-'Ago'
```

```{r}
Trips_01locs<-GSM_locs %>%
  filter(month==This_month)
```

### Step 2: plot_check

```{r}
my_central_location<-data.frame(Longitude=-110.33979846296234,Latitude=24.28728834326802)
```

```{r}
plot_check<-function(my_locs=my_locs,my_central_location=my_central_location){
  ggplot()+
  geom_point(data = my_locs, aes(x=Longitude, y = Latitude),
             color='black',size = 0.8,alpha=0.4)+
  geom_point(data=my_central_location, aes(x=Longitude, y=Latitude),color='red',shape=17, size=5)+
  geom_density_2d_filled(data = my_locs, aes(x = Longitude, y = Latitude),alpha = 0.5)+
  theme_bw()+
  theme(legend.position = 'none')+
  ggtitle('Check that the triangle is close to the highest location concentration- \nThis should be your central location-  \nIf the central location is shifted -  \nPlease recalculate using another central location')}
```

```{r}
plot_check(my_locs=Trips_01locs,my_central_location = my_central_location)
```

### Step 3: remove central locations

```{r}
Trips_02outside<-Trips_01locs %>%
  filter(inside=='outside_central')
```

### Step 4: trip_number_sequence

```{r,echo=FALSE, eval=FALSE}
usethis::use_r("continue_trip_sequence")

# ROXIGEN fill up
#' A function to calculate gaps between locations inside trips
#'
#' @param GPS_data the data containing date, time and trip number
#' @param column_datetime a column with date and time information already in POSIXct format
#' @param column_tripnumber a column with trip_number or any other separator
#'
#' @return the same data frame with an additional column called timedif_min
#' @export
#'
#' @examples Previous_params<-data.frame(trip_id=c("trip_00405"))
#' trip_number_sequence<-continue_trip_sequence(my_previous=Previous_params$trip_id)
#' 
continue_trip_sequence<-function(my_previous){
  print(dplyr::last(my_previous))
  trip_number_sequence<-substr(dplyr::last(my_previous), start = 6, stop = 10)
  trip_number_sequence<-as.numeric(trip_number_sequence)
  print(trip_number_sequence)
return(trip_number_sequence)}

# Run in console
devtools::document()
usethis::use_package("dplyr")
devtools::check()
```

```{r}
Previous_params<-data.frame(trip_id=c("trip_00405"))
```

```{r}
trip_number_sequence<-continue_trip_sequence(my_previous=Previous_params$trip_id)
```

```{r}
Trips_03trips<-Trips_02outside %>%
  mutate(num_seq=as.numeric(num_seq))%>%
  mutate(trip_number = (cumsum(c(1L, diff(num_seq)) !=   1L)))%>%
  mutate(trip_number = trip_number +1 + trip_number_sequence)%>%
  mutate(trip_number = stringr::str_pad(trip_number,  5, pad = "0"))%>%
  mutate(trip_number = paste0("trip_", trip_number))
```

### Step 5: plot_trips

function: plot_trips

```{r}
plot_trips<-function(my_locs=my_locs,my_central_location=my_central_location){
  ggplot()+
  geom_point(data = my_locs, aes(x=Longitude, y = Latitude,color=trip_number),
            size = 0.8,alpha=0.4)+
  geom_point(data=my_central_location, aes(x=Longitude, y=Latitude),color='red',shape=17, size=5)+
  geom_density_2d_filled(data = my_locs, aes(x = Longitude, y = Latitude),alpha = 0.1)+
  theme_bw()+
  theme(legend.position = 'none')+
  ggtitle('Here excludes all nest locations \nCheck if there is any pattern that suggest that the central location is shifted \nThis would include hotstops (brighter colors) in a specific area')}
```

```{r}
plot_trips(my_locs=Trips_03trips,my_central_location=my_central_location)
```

### Step 6: check parameters

```{r}
library(sula)
```

```{r}
Trips_04params<-sula::calcular_duracion(GPS_data=Trips_03trips, 
                        col_diahora='daytime',
                        formato =  "%Y-%m-%d %H:%M:%S",
                        unidades="hours",
                        separador="trip_number")
```

Define trip_month_id in case there are outliers and the locations need to be subset

```{r}
Trips_04params<-Trips_04params %>%
  mutate(trip_month_id=paste0(This_month_text,"_a_",trip_id))%>%
  mutate(central_loc='colony')
```

### Step 7: gaps_params

```{r}
gaps_params<-function(my_params=my_params,my_locs=my_locs){
  
  gaps_params<-my_locs %>%
    group_by(trip_number)%>%
    summarise(min_gap=min(gaps_time),
              max_gap=max(gaps_time))%>%
    rename(trip_id=trip_number)

    this_params<-gaps_params %>%
      left_join(my_params,by='trip_id')%>%
      select(trip_id,trip_month_id,central_loc,trip_start,trip_end,duration,min_gap,max_gap)
  
  return(this_params)
}
```

```{r}
Trips_05params<-gaps_params(my_params = Trips_04params, my_locs = Trips_03trips)
head(Trips_05params)
```


# Short: straight forward using functions

```{r}
Trips_01locs<-my_locs%>%
  filter(month==This_month)
```

```{r}
plot_check(my_locs=Trips_01locs,my_central_location=my_central_location)
```

```{r}
Trips_02outside<-Trips_01locs %>%
  filter(inside=='outside_central')
```

```{r}
trip_number_sequence<-continue_trip_sequence(my_previous=Trips_05params$trip_id)
```

```{r}
Trips_03trips<-Trips_02outside %>%
  mutate(num_seq=as.numeric(num_seq))%>%
  mutate(trip_number = (cumsum(c(1L, diff(num_seq)) !=   1L)))%>%
  mutate(trip_number = trip_number +1 + trip_number_sequence)%>%
  mutate(trip_number = stringr::str_pad(trip_number,  5, pad = "0"))%>%
  mutate(trip_number = paste0("trip_", trip_number))
```

```{r}
plot_trips(my_locs=Trips_03trips,my_central_location=my_central_location)
```

```{r}
Trips_04params<-sula::calcular_duracion(GPS_data=Trips_03trips, 
                        col_diahora='daytime',
                        formato =  "%Y-%m-%d %H:%M:%S",
                        unidades="hours",
                        separador="trip_number")
```

```{r}
range(Trips_04params$duration)
```

```{r}
hist(Trips_04params$duration)
```

```{r}
Trips_04params<-Trips_04params %>%
  mutate(trip_month_id=paste0(This_month_text,"_a_",trip_id))%>%
  mutate(central_loc='colony')
```

```{r}
Trips_05params<-gaps_params(my_params = Trips_04params, my_locs = Trips_03trips)
head(Trips_05params)
```


```{r}
length(unique(Trips_05params$trip_id))
length(unique(Trips_03trips$trip_number))
```


# ~~~~~~~~~~~~~~~~~~~~~~~~

# Check criteria

## params_class

```{r}
params_class<-function(my_params=my_params){
  my_params<-my_params %>%
    mutate(trip_size = case_when(duration >= 0.5 ~ 'trip_longer_than_30mins',
                                 duration <= 0.5 ~ 'trip_shorter_than_30mins',
                                 TRUE ~ "other"))%>%
    
    mutate(resolution=case_when(max_gap >= 60 ~ 'low_resolution_gaps_more_60_mins',
                                max_gap <= 60 ~ 'ok_resolution_gaps_less_60_mins',
                                TRUE ~ "other"))%>%
    
    mutate(params_analyses=case_when(duration >= 24 ~ 'reevaluate_centralloc_longer_than_24h',
                                     duration <= 24 ~ 'keep_centralloc_shorter_than_24h',
                                     TRUE ~ "other"))%>%
    
    mutate(interpolation = case_when(duration <= 24 & max_gap <= 60 ~ 'interpolate_gapsless60mins_shorter24hr',
                                     duration >= 24 & max_gap >= 60 ~ 'dont_interpolate_gapsmore60mins_longer24hr',
                                     duration >= 24 ~ 'dont_interpolate_longer24hr',
                                     max_gap >= 60 ~ 'dont_interpolate_gapsmore60mins_longer24hr',
                                     TRUE ~ "other"))
  
  return(my_params)}
```

```{r}
Params_00criteria<-params_class(my_params=Trips_05params)
```

## check_trip_criteria

Foraging trips were only considered when the animal was more than 1 km away from the colony and lasting longer than 30 min (Shaffer et al. 2017, Guerra et al. 2022). 

If trips had a gap of > 1hr and/or included overnight locations without a clear central location, they were excluded from the interpolation.

```{r}
check_trip_criteria<-function(my_params=my_params){
  
  trip_criteria<-my_params %>%
    group_by(trip_size)%>%
    tally()%>%
    mutate(total=sum(n))%>%
    mutate(prop=n*100/total)
  
   print(paste0("From ",sum(trip_criteria$n),' trips: ',
                round(trip_criteria$prop[1],2),' were ',trip_criteria$trip_size[1],' and ',
                 round(trip_criteria$prop[2],2),' were ',trip_criteria$trip_size[2],
                '. Trips shorter than 30 minutes are not considered real trips. Remove trips shorter than 30 minutes'
                ))
  
}
```

```{r}
check_trip_criteria(my_params=Params_00criteria)
```
```{r}
Params_01criteria<-Params_00criteria %>%
  filter(trip_size == 'trip_longer_than_30mins')
```

## check_resolution_criteria

```{r}
check_resolution_criteria<-function(my_params=my_params){
  
  resolution_criteria<-my_params %>%
    group_by(resolution)%>%
    tally()%>%
    mutate(total=sum(n))%>%
    mutate(prop=n*100/total)

  
  print(paste0("From ",sum(resolution_criteria$n),' trips: ',
                round(resolution_criteria$prop[1],2),' (n = ',resolution_criteria$n[1],') were ',
                                                                  resolution_criteria$resolution[1],' and ',
                 round(resolution_criteria$prop[2],2),' (n = ',resolution_criteria$n[2],') were ',
               resolution_criteria$resolution[2],
                '. Evaluate if trips with low resolution are to be kept'))
  
}
```

```{r}
check_resolution_criteria(my_params=Params_01criteria)
```

## check_lenght_criteria

```{r}
Params_02criteria<-Params_01criteria %>%
  filter(resolution == 'ok_resolution_gaps_less_60_mins')
```

```{r}
check_lenght_criteria<-function(my_params=my_params){
  
  lenght_criteria<-my_params %>%
    group_by(params_analyses)%>%
    tally()%>%
    mutate(total=sum(n))%>%
    mutate(prop=n*100/total)

  
  print(paste0("From ",sum(lenght_criteria$n),' trips: ',
                round(lenght_criteria$prop[1],2),' (n = ',lenght_criteria$n[1],') were ',lenght_criteria$params_analyses[1],' and ',
                 round(lenght_criteria$prop[2],2),' (n = ',lenght_criteria$n[2],') were ',lenght_criteria$params_analyses[2],
                '. Evaluate if trips longer than 24 hrs is because of a change in central location'))
  
}
```

```{r}
check_lenght_criteria(my_params=Params_02criteria)
```

# Reevaluate

```{r}
Reevaluate<-Params_02criteria %>%
  filter(params_analyses=='reevaluate_centralloc_longer_than_24h')
Reevaluate
```

# ~~~~~~~~~~~~~~~~~

# 3. Gulls just do whatever they want

## Step 1: check outliers

Check big outliers

```{r}
Reevaluate<-Params_02criteria %>%
  filter(params_analyses=='reevaluate_centralloc_longer_than_24h')
Reevaluate
```

Are this outliers due to gaps of the intervals?
Large intervals might cause that the bird entering back to the colony is missed.

```{r}
print(paste0("Max gaps was: ",
             round(Reevaluate$max_gap,2),
             ' mins. If the interval is more than 60 minutes in duration that could be a plausible explaination and then it should be eliminated. If the interval is less than 60 minutes in duration consider it can be a change in central location of the bird.'))
```

## Step 2: extract locs

Identify ids 

```{r}
Reevaluate_tripid<-unique(Reevaluate$trip_id)
Reevaluate_tripid
```


## Step 3.Subset locations

```{r}
Reevaluate_01locs<-Trips_03trips %>%
  filter(trip_number %in% Reevaluate_tripid)%>%
  relocate(ID,trip_number,daytime,dt,gaps_time)
head(Reevaluate_01locs)
```

plot_check with facet_wrap

```{r}
plot_check(my_locs=Reevaluate_01locs,my_central_location = my_central_location)+
  facet_wrap(~trip_number)
```


## Step 4: identify new central location

```{r}
new_central_location<-data.frame(Longitude=-110.325,Latitude=24.17)
```

```{r}
plot_check(my_locs=Reevaluate_01locs,my_central_location = my_central_location)+
  geom_point(data=new_central_location, aes(x=Longitude, y=Latitude),color='blue',shape=17, size=5)
```



## Step 5: create_buffer

```{r}
create_buffer<-function(central_point=central_point, buffer_km=buffer_km){
  central_spatial<- sp::SpatialPoints(cbind(central_point$Longitude,central_point$Latitude)) 
  sp::proj4string(central_spatial)= sp::CRS("+init=epsg:4326") 
  central_spatial <- sp::spTransform(central_spatial, sp::CRS("+init=epsg:4326"))
  central_spatial<-sf::st_as_sf(central_spatial)
  buffer_dist<-buffer_km*1000
  central_buffer<-sf::st_buffer(central_spatial, buffer_dist)
  return(central_buffer)
  }
```

```{r}
new_central_buffer<-create_buffer(central_point=new_central_location,buffer_km=0.3)
```

## Step 6: from_df_to_sf

```{r}
from_df_to_st<-function(my_df){
  my_points <- my_df
  sp::coordinates(my_points) <- ~Longitude + Latitude
  sp::proj4string(my_points) = sp::CRS("+init=epsg:4326")
  my_sf<-sf::st_as_sf(my_points)
  return(my_sf)
}
```

```{r}
Reevaluate_02sf<-from_df_to_st(Reevaluate_01locs)
```

## Step 7: over

```{r}
over<-function(my_locations=my_locations,my_buffer=my_buffer){
  locations<-sf::st_as_sf(my_locations)
  location_over<-sapply(sf::st_intersects(locations,my_buffer), 
                        function(z) if (length(z)==0) NA_integer_ else z[1])
  return(location_over)}
```

```{r}
Reevaluate_03over<- over(my_locations=Reevaluate_02sf,my_buffer=new_central_buffer)
```

## Step 8: add_trip_number

```{r}
add_trip_number<-function(my_over,my_df,my_trip_number_sequence){
  
  my_df$central_location <- my_over 
  
  my_df<-my_df %>%
    mutate(inside=case_when(central_location == 1 ~ 'inside_central',TRUE ~ 'outside_central'))
  
  my_df %>%
    group_by(ID,inside)%>%
    count()%>%
    pivot_wider(names_from = inside, values_from = n)
  
  my_outside<-my_df %>%
    filter(inside=='outside_central')
  
  my_trips<-my_outside %>%
    mutate(num_seq=as.numeric(num_seq))%>%
    mutate(trip_number = (cumsum(c(1L, diff(num_seq)) !=   1L)))%>%
    mutate(trip_number = trip_number + 1 + my_trip_number_sequence)%>%
    mutate(trip_number = stringr::str_pad(trip_number,  5, pad = "0"))%>%
    mutate(trip_number = paste0("trip_", trip_number))
  
  return(my_trips)
}
```

```{r}
new_trip_number_sequence<-continue_trip_sequence(my_previous=Trips_05params$trip_id)
```

```{r}
Reevaluate_04trips<-add_trip_number(my_df=Reevaluate_01locs,
                                     my_over=Reevaluate_03over,
                                     my_trip_number_sequence=new_trip_number_sequence)
```

## Step 9: calculate params

```{r}
Reevaluate_05params<-sula::calcular_duracion(GPS_data=Reevaluate_04trips, 
                        col_diahora='daytime',
                        formato =  "%Y-%m-%d %H:%M:%S",
                        unidades="hours",
                        separador="trip_number")
```

```{r}
Reevaluate_05params<-Reevaluate_05params %>%
  mutate(trip_month_id=paste0(This_month_text,"_b_",trip_id))%>%
  mutate(central_loc='south_of_colony')
```

## Step 10: gaps_params

```{r}
Reevaluate_06params<-gaps_params(my_params = Reevaluate_05params, my_locs = Reevaluate_04trips)
Reevaluate_06params
```

### checks

```{r}
hist(Reevaluate$duration)
```

```{r}
hist(Reevaluate_06params$duration)
```

## Step 11: plot_trips

```{r}
plot_trips(my_locs=Reevaluate_04trips,my_central_location = my_central_location)+
  geom_point(data=new_central_location, aes(x=Longitude, y=Latitude),color='blue',shape=17, size=5)
```


## Step 12: Params merged

Add criteria classifications

```{r}
Params_03reevaluate<-params_class(my_params=Reevaluate_06params)
```

```{r}
Params_02criteria
Params_03reevaluate
```

```{r}
Params_merged<-rbind(Params_02criteria %>% filter(!trip_id %in% Reevaluate_tripid),
                     Params_03reevaluate)
```

## Step 13: Locs merged

```{r}
Locs_merged <-rbind(Trips_03trips%>% 
                      filter(!trip_number %in% Reevaluate_tripid)%>%
                      select(ID,daytime,season,month,
                             Longitude,Latitude,countday,
                             Hour,Day_or_night,central_base,inside,num_seq,
                             ID.y,dt,Location_gaps,Battery,Gaps_time,battery.charge.percent,
                             trip_number),
                    
                    Reevaluate_04trips %>%
                      select(ID,daytime,season,month,
                             Longitude,Latitude,countday,
                             Hour,Day_or_night,central_base,inside,num_seq,
                             ID.y,dt,Location_gaps,Battery,Gaps_time,battery.charge.percent,
                             trip_number))
```

# ~~~~~~~~~~~~~~~~

# Interpolation

```{r}
Inter_params<-Params_merged
Inter_locs<-Locs_merged
```

## Step 1: exclude trips longer than 24 hrs with gaps larger than 60 minutes

```{r}
Inter_params %>%
  group_by(interpolation)%>%
  tally()
```

```{r}
Inter_01params<-Inter_params %>%
  filter(interpolation=='interpolate_gapsless60mins_shorter24hr')
```

## Step 2: locs_class

```{r}
locs_class<-function(Inter_params=Inter_params,
                     Inter_locs=Inter_locs){
  
  tripid_01trip_size<-unique(Inter_params %>%
                            filter(trip_size == 'trip_longer_than_30mins'))$trip_id
  
  tripid_02analysesid<-unique(Inter_params %>%
                                filter(params_analyses == "keep_centralloc_shorter_than_24h"))$trip_id
  
  tripid_03resolutionid<-unique(Inter_params %>%
                                  filter(resolution == "ok_resolution_gaps_less_60_mins"))$trip_id
  
  tripid_04interpolateid<-unique(Inter_params %>%
                                   filter(interpolation == 'interpolate_gapsless60mins_shorter24hr'))$trip_id
  
  Inter_locs<-Inter_locs %>%
  mutate(trip_size = case_when(trip_number %in% tripid_01trip_size ~  'trip_longer_than_30mins',
                                 TRUE ~  'trip_shorter_than_30mins'))%>%
  
  mutate(params_analyses=case_when(trip_number %in% tripid_02analysesid ~ "keep_centralloc_shorter_than_24h",
                          TRUE ~ "reevaluate_centralloc_longer_than_24h"))%>%
  
  mutate(resolution=case_when(trip_number %in% tripid_03resolutionid~ "ok_resolution_gaps_less_60_mins",
                          TRUE ~ "low_resolution_gaps_more_60_mins"))%>%
  
  mutate(interpolation=case_when(trip_number %in% tripid_04interpolateid ~ 'interpolate_gapsless60mins_shorter24hr',
                          TRUE ~ "dont_interpolate_longer24hr"))
  
  return(Inter_locs)
}
```


```{r}
Inter_02locs<-locs_class(Inter_params=Inter_params,
                     Inter_locs=Inter_locs)
```

## Step 3: check locations

```{r}
Inter_02locs %>%
  group_by(interpolation)%>%
  tally()
```

```{r}
plot_trips(my_locs=Inter_02locs,my_central_location = my_central_location)
```

### Included

```{r}
Inter_04locs <- Inter_02locs %>%
  filter(interpolation == 'interpolate_gapsless60mins_shorter24hr')
```

```{r}
plot_trips(my_locs=Inter_04locs,my_central_location = my_central_location)
```

### Excluded

```{r}
Inter_03locs <- Inter_02locs %>%
  filter(interpolation != 'interpolate_gapsless60mins_shorter24hr')
```

```{r, eval=FALSE, fig.height=20, fig.width=20}
plot_trips(my_locs=Inter_03locs,my_central_location = my_central_location)+
  facet_wrap(~trip_number)
```


Last 6 trips look genuinely that the bird change location

# X
# X
# X

# Reevaluate

```{r}
reevaluate_locs<-function(this_tripid,
                     this_locs,
                     this_central_location){
  
## Step 2: extract locs
Reevaluate_00tripid<-this_tripid
Reevaluate_00tripid

## Step 3.Subset locations
Reevaluate_01locs<-this_locs %>%
  filter(trip_number %in% Reevaluate_00tripid)%>%
  relocate(ID,trip_number,daytime,dt,gaps_time)

new_central_location<-this_central_location

plot_check(my_locs=Reevaluate_01locs,my_central_location = this_central_location)+
  geom_point(data=new_central_location, aes(x=Longitude, y=Latitude),color='blue',shape=17, size=5)

return(Reevaluate_01locs)}
```

```{r}
revaluate_params<-function(this_tripid,
                     this_locs,
                     this_central_location,
                     this_previous){
  
## Step 5: create_buffer
new_central_buffer<-create_buffer(central_point=new_central_location,buffer_km=0.3)

## Step 6: from_df_to_sf
Reevaluate_02sf<-from_df_to_st(Reevaluate_01locs)

## Step 7: over
Reevaluate_03over<- over(my_locations=Reevaluate_02sf,my_buffer=new_central_buffer)

## Step 8: add_trip_number
new_trip_number_sequence<-continue_trip_sequence(my_previous=this_previous$trip_id)

Reevaluate_04trips<-add_trip_number(my_df=Reevaluate_01locs,
                                     my_over=Reevaluate_03over,
                                     my_trip_number_sequence=new_trip_number_sequence)
## Step 9: calculate params
Reevaluate_05params<-sula::calcular_duracion(GPS_data=Reevaluate_04trips, 
                        col_diahora='daytime',
                        formato =  "%Y-%m-%d %H:%M:%S",
                        unidades="hours",
                        separador="trip_number")

Reevaluate_05params<-Reevaluate_05params %>%
  mutate(trip_month_id=paste0(This_month_text,"_b_",trip_id))%>%
  mutate(central_loc='south_of_colony')

## Step 10: gaps_params
Reevaluate_06params<-gaps_params(my_params = Reevaluate_05params, my_locs = Reevaluate_04trips)
Reevaluate_06params

return(Reevaluate_06params)}
```

### checks

```{r}
hist(Reevaluate$duration)
```

```{r}
hist(Reevaluate_06params$duration)
```

## plot_trips

```{r}
plot_trips(my_locs=Reevaluate_04trips,my_central_location = my_central_location)+
  geom_point(data=new_central_location, aes(x=Longitude, y=Latitude),color='blue',shape=17, size=5)
```


Whether to interpolate the data depends on your research goals and the nature of the gaps.

Here’s a breakdown of considerations to help you decide:

- Short Gaps: If the gaps are short (e.g., a few minutes to an hour), interpolation might be justifiable, especially if you assume that the bird's movement is relatively constant during the gap. For example, linear interpolation can estimate a reasonable position based on previous and subsequent data points.
- Long Gaps: If the gaps are long (e.g., several hours or days), interpolation might introduce unrealistic results, especially if the bird’s movement during that time period was unknown. Interpolating long gaps could lead to misleading conclusions about the bird’s actual path or behavior.
If your study’s analysis can handle missing data or gaps (e.g., for high-level migration patterns), you might consider leaving the gaps unfilled. This avoids the risk of introducing erroneous data but may reduce the overall precision.
If your analysis is based on trajectory or movement behavior, you could use gap-filling approaches like predicting the bird’s behavior based on its last known location and surrounding environmental context (such as weather, time of day, etc.).




# ~~~~~~~~~~~~~~~~~~~~~
# Test data

## gull_data

Contains data from one gull tagged at La Paz, Mexico

```{r,echo=FALSE,eval=FALSE}
#load packages
library(tidyverse)
#load data
LALIS_1807 <- read_csv("~/03Academico/01MS/01EnPreparacion/MS4 Lali/2023/02Jul/1Data/LALIS_1807.csv")
range(LALIS_1807$timestamp)
#rename columns
LALIS<-LALIS_1807 %>%
  filter(timestamp > '2023-04-20 00:00:00 UTC' & timestamp > '2023-04-27 00:00:00 UTC')
LALIS<-LALIS %>%
  select(timestamp,location.long,location.lat,individual.local.identifier)
LALIS$Longitude<-LALIS$location.long
LALIS$Latitude<-LALIS$location.lat
#identify trips
LALI02<-LALIS%>%
  filter(individual.local.identifier=='LALI02')
#devtools::install_github("MiriamLL/sula")
library(sula)
colony<-data.frame(Longitude=-110.338935,Latitude=24.287935)
LALIS_trips<-identificar_viajes(GPS_data=LALI02,
                        nest_loc=colony,
                        distancia_km=1)
LALIS_trips$at_sea<-LALIS_trips$trip
LALIS_atsea<-contar_viajes(GPS_data=LALIS_trips)
gull_data<-LALIS_atsea %>%
  select(individual.local.identifier,
         Longitude,Latitude,
         timestamp,
         at_sea,
         trip_number)%>%
  rename(ID=individual.local.identifier,
         DateTime=timestamp)
ggplot()+
  geom_point(data=gull_data, aes(x=Longitude,y=Latitude,color=trip_number))+
  theme_minimal()+
  theme(legend.position = 'none')
gull_data<-gull_data[1:1458,]
save(gull_data, file = "gull_data.RData", version = 2)
#usethis::use_r("gull_data")
devtools::document()
devtools::check()
```

```{r}
library(larus)
```

```{r}
gull_data<-gull_data
```

