---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# larus

<!-- badges: start -->
<!-- badges: end -->

The goal of larus is to 

Provide data example

Identify foraging trips from a central location
Using:
a buffer
converting dt to st
using over 
add a trip number
keep the sequence

Identify is trips are valid by:
calculating gaps between locations
evaluate if trips are valid by checking thresholds (e.g. minimum duration away from the colony)
lenght paths requeire at least 3 locations per forafging trip
exclude trips where the bird was constantly moving from central location

recalculate foraging trips changing central locations

Interpolate without:
overestimating resting locations at night (make a filter of trips <24 hrs)
add a classification trip good for calculating foraging parameters
trip good for interpolation 


## Installation

You can install the development version of larus from [GitHub](https://github.com/) with:

```{r,eval=FALSE}
# install.packages("devtools")
devtools::install_github("MiriamLL/larus")
```

# Main functions

## Check for gaps and resolution

- from_df_to_st
- over
- identify_trips 
- trip_number_sequence
- calculate_gaps

# Data

## gull_data

Contains data from one gull tagged at La Paz, Mexico

```{r,echo=FALSE,eval=FALSE}
#load packages
library(tidyverse)
#load data
LALIS_1807 <- read_csv("~/03Academico/01MS/01EnPreparacion/MS4 Lali/2023/02Jul/1Data/LALIS_1807.csv")
range(LALIS_1807$timestamp)
#rename columns
LALIS<-LALIS_1807 %>%
  filter(timestamp > '2023-04-20 00:00:00 UTC' & timestamp > '2023-04-27 00:00:00 UTC')
LALIS<-LALIS %>%
  select(timestamp,location.long,location.lat,individual.local.identifier)
LALIS$Longitude<-LALIS$location.long
LALIS$Latitude<-LALIS$location.lat
#identify trips
LALI02<-LALIS%>%
  filter(individual.local.identifier=='LALI02')
#devtools::install_github("MiriamLL/sula")
library(sula)
colony<-data.frame(Longitude=-110.338935,Latitude=24.287935)
LALIS_trips<-identificar_viajes(GPS_data=LALI02,
                        nest_loc=colony,
                        distancia_km=1)
LALIS_trips$at_sea<-LALIS_trips$trip
LALIS_atsea<-contar_viajes(GPS_data=LALIS_trips)
gull_data<-LALIS_atsea %>%
  select(individual.local.identifier,
         Longitude,Latitude,
         timestamp,
         at_sea,
         trip_number)%>%
  rename(ID=individual.local.identifier,
         DateTime=timestamp)
ggplot()+
  geom_point(data=gull_data, aes(x=Longitude,y=Latitude,color=trip_number))+
  theme_minimal()+
  theme(legend.position = 'none')
gull_data<-gull_data[1:1458,]
save(gull_data, file = "gull_data.RData", version = 2)
#usethis::use_r("gull_data")
devtools::document()
devtools::check()
```

```{r}
library(larus)
```

```{r}
gull_data<-gull_data
```


# Functions

## create_buffer

```{r}
create_buffer<-function(central_point=central_point, buffer_km=buffer_km){
  central_spatial<- sp::SpatialPoints(cbind(central_point$Longitude,central_point$Latitude)) 
  sp::proj4string(central_spatial)= sp::CRS("+init=epsg:4326") 
  central_spatial <- sp::spTransform(central_spatial, sp::CRS("+init=epsg:4326"))
  central_spatial<-sf::st_as_sf(central_spatial)
  buffer_dist<-buffer_km*1000
  central_buffer<-sf::st_buffer(central_spatial, buffer_dist)
  return(central_buffer)
  }
```

```{r,eval=FALSE}
central_buffer<-create_buffer(central_point=central_location,buffer_km=0.3)
```

## from_df_to_st

```{r,eval=FALSE}
LALI02_02apr_points <- LALI02_02apr
sp::coordinates(LALI02_02apr_points) <- ~Longitude + Latitude
sp::proj4string(LALI02_02apr_points) = sp::CRS("+init=epsg:4326")
LALI02_02apr_points<-sf::st_as_sf(LALI02_02apr_points)
```

```{r}
from_df_to_st<-function(my_df){
  my_points <- my_df
  sp::coordinates(my_points) <- ~Longitude + Latitude
  sp::proj4string(my_points) = sp::CRS("+init=epsg:4326")
  my_sf<-sf::st_as_sf(my_points)
  return(my_sf)
}
```

```{r,eval=FALSE}
LALI02_27ago_points<-from_df_to_st(LALI02_27ago)
```

## over

```{r,eval=FALSE}
#LALI02_02apr_over<- over(LALI02_02apr_points,central_buffer)
LALI02_02apr_over<-sapply(sf::st_intersects(LALI02_02apr_points,central_buffer), function(z) if (length(z)==0) NA_integer_ else z[1])
```

```{r,eval=FALSE}
LALI02_02apr$central_location <- LALI02_02apr_over
```

```{r}
over<-function(this_location=this_location,this_buffer=this_buffer){
  locations<-sf::st_as_sf(this_location)
  location_over<-sapply(sf::st_intersects(locations,this_buffer), function(z) if (length(z)==0) NA_integer_ else z[1])
  return(location_over)}
```

```{r,eval=FALSE}
LALI02_27ago_over<- over(LALI02_27ago_points,central_buffer)
```

## count_trip

```{r,eval=FALSE}
LALI02_02apr<-LALI02_02apr %>%
  mutate(inside=case_when(central_location == 1 ~ 'inside_central',TRUE ~ 'outside_central'))
```

```{r,eval=FALSE}
LALI02_02apr %>%
  group_by(ID,inside)%>%
  count()%>%
  pivot_wider(names_from = inside, values_from = n)
```

```{r,eval=FALSE}
LALI02_04apr<-LALI02_02apr %>%
  filter(inside=='outside_central')
```

```{r,eval=FALSE}
LALI02_04apr<-LALI02_04apr %>%
  mutate(num_seq=as.numeric(num_seq))%>%
  mutate(trip_number = (cumsum(c(1L, diff(num_seq)) !=   1L)))%>%
  mutate(trip_number = trip_number + 1)%>%
  mutate(trip_number = stringr::str_pad(trip_number,  5, pad = "0"))%>%
  mutate(trip_number = paste0("trip_", trip_number))
```

## trip_number_sequence

```{r,eval=FALSE}
last(LALI02_04apr$trip_number)
```

```{r,eval=FALSE}
trip_number_sequence<-substr(last(LALI02_04apr$trip_id), start = 6, stop = 10)
trip_number_sequence<-as.numeric(trip_number_sequence)
trip_number_sequence
```

```{r}
continue_trip_sequence<-function(my_previous){
  print(last(my_previous$trip_id))
  trip_number_sequence<-substr(last(my_previous$trip_id), start = 6, stop = 10)
  trip_number_sequence<-as.numeric(trip_number_sequence)
  print(trip_number_sequence)
return(trip_number_sequence)}
```

```{r,eval=FALSE}
trip_number_sequence<-continue_trip_sequence(my_previous=LALI02_54oct)
```

### add_trip_number

```{r}
add_trip_number<-function(my_over,my_df,my_previous){
  
  my_df$central_location <- my_over 
  
  my_df<-my_df %>%
    mutate(inside=case_when(central_location == 1 ~ 'inside_central',TRUE ~ 'outside_central'))
  
  my_df %>%
    group_by(ID,inside)%>%
    count()%>%
    pivot_wider(names_from = inside, values_from = n)
  
  my_outside<-my_df %>%
    filter(inside=='outside_central')
  
  last(my_previous$trip_id)
  
  trip_number_sequence<-substr(last(my_previous$trip_id), start = 6, stop = 10)
  trip_number_sequence<-as.numeric(trip_number_sequence)
  trip_number_sequence

  my_trips<-my_outside %>%
    mutate(num_seq=as.numeric(num_seq))%>%
    mutate(trip_number = (cumsum(c(1L, diff(num_seq)) !=   1L)))%>%
    mutate(trip_number = trip_number + 1 + trip_number_sequence)%>%
    mutate(trip_number = stringr::str_pad(trip_number,  5, pad = "0"))%>%
    mutate(trip_number = paste0("trip_", trip_number))
  
  return(my_trips)
}
```

```{r,eval=FALSE}
LALI02_28ago<-add_trip_number(my_over=LALI02_27ago_over,my_df=LALI02_27ago,my_previous=LALI02_13jul)
```


### sequence

```{r,eval=FALSE}
LALI02_27ago_points<-from_df_to_st(LALI02_27ago)
```

```{r,eval=FALSE}
LALI02_27ago_over<- over(LALI02_27ago_points,central_buffer)
```

```{r,eval=FALSE}
trip_number_sequence<-continue_trip_sequence(my_previous=LALI02_54oct$trip_id)
```

```{r,eval=FALSE}
LALI02_28ago<-add_trip_number(my_over=LALI02_27ago_over,my_df=LALI02_27ago,my_previous=LALI02_13jul)
```

# Gaps

```{r,eval=FALSE}
LALIS_all$Gaps_time<-as.numeric(LALIS_all$daytime - lag(LALIS_all$daytime))
```

```{r, echo=FALSE, eval=FALSE}
calculate_gaps<-function(GPS_data=GPS_data,
                         column_datetime=column_datetime,
                         column_tripnumber=column_tripnumber){
  
  GPS_data$column_tripnumber<-(GPS_data[[column_tripnumber]])
  
  trips_list<-split(GPS_data,GPS_data$column_tripnumber)
  
  gaps_list<-list()
  
  for( i in seq_along(trips_list)){
    
    trip_df<-trips_list[[i]]

    times<-trip_df[[column_datetime]]
    times_lag<-lag(times)
    time_dif<-as.numeric(difftime(times,times_lag, units="mins"))
    trip_df$timedif_min<-round(time_dif,2)
    
  gaps_list[[i]]<-trip_df
  
  }
  
  gaps_df<- do.call("rbind",gaps_list)
  
  gaps_df$column_tripnumber<-NULL

  return(gaps_df)
}
```

```{r}
gull_gaps<-calculate_gaps(GPS_data = gull_data,
                          column_datetime = 'DateTime',
                          column_tripnumber = 'trip_number')
```

```{r}
range(gull_gaps$timedif_min,na.rm=TRUE)
```


# Intervals

When studying tracking locations from GSM data, especially in the case of birds like gulls, it’s common to encounter gaps in the data due to factors such as loss of signal, battery limitations, or movement through areas with poor GSM coverage. Whether to interpolate the data depends on your research goals and the nature of the gaps.

Here’s a breakdown of considerations to help you decide:

- Short Gaps: If the gaps are short (e.g., a few minutes to an hour), interpolation might be justifiable, especially if you assume that the bird's movement is relatively constant during the gap. For example, linear interpolation can estimate a reasonable position based on previous and subsequent data points.
- Long Gaps: If the gaps are long (e.g., several hours or days), interpolation might introduce unrealistic results, especially if the bird’s movement during that time period was unknown. Interpolating long gaps could lead to misleading conclusions about the bird’s actual path or behavior.
If your study’s analysis can handle missing data or gaps (e.g., for high-level migration patterns), you might consider leaving the gaps unfilled. This avoids the risk of introducing erroneous data but may reduce the overall precision.
If your analysis is based on trajectory or movement behavior, you could use gap-filling approaches like predicting the bird’s behavior based on its last known location and surrounding environmental context (such as weather, time of day, etc.).

# Battery

```{r}
library(scales)
```

```{r,eval=FALSE}
ggplot(LALI02_01locs, aes(x=daytime, y=battery.charge.percent)) +
  geom_line() +
  scale_x_datetime(labels = date_format("%b"),date_breaks = "1 month")
```

# Classify



Based on the trip parameters, add a column in the locations if the trip would be interpolated or not.

Foraging trips were only considered when lasting longer than 30 min 

```{r,eval=FALSE}
LALI02_threshold_id<<-unique(LALI02_04params %>%
  filter(duration > 0.5))$trip_id
```

```{r,eval=FALSE}
LALI02_04locs_trips<- LALI02_03locs %>%
  filter(trip_number %in% LALI02_threshold_id)
```

If trips are longer than 1 hr, then the parameters are calculable.

```{r,eval=FALSE}
LALI02_04locs_trips$params<-'calculable'
```

